### 2024-10-01 MetadataS1/MetadataS1.tex
update table variable definition and complete first draft of the entire document

\documentclass[a4paper,twoside,12pt]{article}

%% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}
\usepackage{lmodern} %Type1-font for non-english texts and characters
\usepackage{parskip}
\usepackage{enumitem}

%% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx} %%For loading graphic files
\usepackage{subcaption}
\usepackage{float}
\usepackage{booktabs} %% for top/midrule
\usepackage{textcomp}
\usepackage[table,xcdraw, dvipsnames]{xcolor}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{makecell}
\usepackage{pdflscape}

%bibliography
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0} % Removes the border
}
\usepackage{geometry}

%% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}

%% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\onehalfspacing       %% 1,5-spacing
\usepackage{placeins}
\usepackage{lineno}
\linenumbers

\usepackage{titlesec}
\titlespacing{\section}{0pt}{*4}{*1.5} 
\titlespacing{\subsection}{20pt}{*4}{*1.5}
\titlespacing{\subsubsection}{40pt}{*4}{*1.5} 
\titlespacing{\paragraph}{60pt}{*4}{*1.5}
\titlespacing{\subparagraph}{80pt}{*4}{*1.5}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begingroup  
  \centering
\textbf{Title: Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]

\textbf{Authors}:\\
 Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1}, 
 Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]

\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{5} Centre d’études nordiques, Centre for Arctic Knowledge and Exploration, Canadian Museum of Nature, Ottawa, ON, Canada\\
\textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
\\[1.5em]

\textbf{Corresponding Author}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
\textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}
\\[1.5em]

\textbf{Open Research statement}:

%%%%%%% !!!! ADD DRYAD and Zenodo adress !!!! %%%%%%%%
Data are publicly available at \url{https://datadryad.org/}.\\[1.5em]
Raw data and codes used to extract the presented data set are publicly available at \url{https://zenodo.org/}.
\endgroup
\newpage

\subsection*{Introduction}
 The composition of ecological communities, defined as the abundance of each species within a given community, is fundamental for understanding patterns and processes in community ecology. Variations in community composition can help to detect spatial patterns linked to environmental variations \citep{kemp1990}, assess temporal trends of different groups of species after disturbances \citep{philippi1998, magurran2007}, and understand food web structures \citep{cohen2003}. Additionally, community composition is essential for modeling the dynamics of ecological communities. Dynamic community modelling allows addressing important issues and questions in ecology, such as: determining the relative strength of top-down versus bottom-up forces in communities \citep{krebs2003,legagneux2014}, assessing the ecological resilience of communities under climate change \citep{griffith2019} and evaluating the cascading effects of invasive species in food web \citep{david2017, goto2020}. Dynamic community modelling can also be applied to address practical challenges, including fishery management (Plagányi, 2007) and the planning of protected areas \citep{okey2004, dahood2020}.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.75\linewidth]{figures/food_web.pdf} 
\caption{Synthetic vertebrate food web of the south plain of Bylot Island.}
\label{figure:food_web}
\end{figure}

\subsection*{Objective}
Our main objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. Our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments. Additionally, we aim to provide the average body mass for each species in the community, enabling the conversion of abundances into biomasses.
\newpage

\section*{\textcolor{Blue}{Class I. Data Set Descriptors}}
    \section*{A. Data set identity} Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut
    \section*{B. Data set identification codes}
     \texttt{BYLOT-species\_taxonomy.csv}\\
     \texttt{BYLOT-species\_abundance.csv}\\
     \texttt{BYLOT-mean\_species\_abundance.csv}\\
     \texttt{BYLOT-mean\_species\_body\_mass.csv}\\
     \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\

    \section*{C. Data set description}
		\subsection*{1. Originators}
		\textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
  		\textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
  		\textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada

		\subsection*{2. Abstract}
Arctic ecosystems present unique opportunities for community-wide monitoring, in part due to their relatively low species richness. However, conducting research in these remote environments poses significant logistical challenges, resulting in long-term monitoring being exceedingly rare. Here, we focus on the long-term, intensive ecological monitoring efforts conducted on the south plain of Bylot Island (almost 400 km\textsuperscript{2}, Nunavut, Canada), which has generated a remarkable dataset spanning up to 30 years, a rarity in tundra ecosystems. Our goal is to synthesize this dataset and upscale vertebrate abundance data at the landscape level, a prerequisite to conduct community-level analyses. We have standardized data obtained with different field methods to provide readily usable long-term time series of abundance for 35 vertebrate species (30 birds and 5 mammals) present in the study system. Monitoring data includes intensive capture-mark-recapture density estimates of lemmings on trapping grids, systematic or opportunistic nest monitoring conducted across the entire study area or within specific plots for all bird species, transects of vertebrate counts distributed throughout the study area, daily incidental observations of vertebrates and satellite tracking of fox movements. Annual abundance of species was estimated at the landscape level, accounting for spatial variations. Furthermore, we provide body masses for each species, derived from empirical onsite measurements for 18 species and from the literature for the remaining species. Body mass is essential to convert species abundance into biomass for studies of trophic fluxes and ecosystem processes. Our dataset provides a unique opportunity for holistic empirical studies of ecological communities, allowing a deeper understanding of community structure and dynamics. Considering that the study site is a pristine and protected area that has experienced minimal anthropogenic impact, it can also provide an ideal baseline for investigating the impacts of global changes on high-latitude terrestrial ecosystems.

   \section*{D. Key words/phrases}
 Bylot Island, Canadian Arctic, Arctic tundra, 1993-2023, long-term monitoring, biodiversity monitoring, community composition, species abundance, species density, species biomass, species body mass, food web
  \newpage

\section*{\textcolor{Blue}{Class II. Research origin descriptors}}
    \section*{A. Overall project description}
    \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial vertebrate communities

	\subsection*{2. Originators:}	
  	\textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
  		\textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
  		\textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada

@@

    \section*{B. Specific subproject description}
     	\subsection*{1. Site description}
 			\subsubsection*{a. Site type}
  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic.An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (\textit{Vulpes vulpes}) was ambiguous. While the presence of breeding pairs has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare. Therefore, we decided to exclude this species.

         		\input{tables/species_name_strategy.tex}
%label-> table:species_name_strategy
\newpage
            	\subsubsection*{b. Geography} Our 463 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
            	\subsubsection*{c. Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
            	\subsubsection*{d. Geology} see \citet{klassen1993} for a complete and detailed description of the geology of the study area.
            	\subsubsection*{e. Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
            	\subsubsection*{f. Site history} see \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
            	\subsubsection*{g. Climate} The mean temperature in July is 6°C, and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}. The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds. \citep{gauthier2024a}.
 \newpage

      	\subsection*{2. Experimental or sampling design}
        		\subsubsection*{a. Permanent plots}
The study area is divided into 9 zones based on the sampling method and the level of field effort applied in each zone (\textbf{Figure \ref{figure:zones}}). Long-term monitoring of the community began in the 1990s in the Qarlikturvik valley \citep{gauthier2013, gauthier2024a}, which represents the zone of the study area with the highest annual sampling effort. Within the Qarlikturvik valley, the sampling is concentrated on the southern side of the glacial river (\textbf{Figure \ref{figure:qarlikturvik_valley}}), where the main research infrastructure is located. Another zone with extensive sampling efforts is Camp 2, located at the core of the snow goose colony, where the primary focus is to monitor snow goose nests. However, nests of many other avian species are also monitored within and around the colony in this zone. Camp 3, Pointe Dufour, Goose Point, and Malaview are zones where intensive sampling efforts are conducted annually, albeit for a relatively brief period (approximately one week) during the breeding season of most species \citep{gauthier2024a}. The upland zones in the study area (defined as areas approximately 300 meters above sea level or more) are the Black Plateau, Southern Plateau, and Camp 3 Plateau. These zones are primarily visited to assess raptor nesting activity \citep{beardsell2016}. The zone between the Qarlikturvik valley and Camp 3 received very little sampling effort and is therefore excluded from the study area.

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth, angle=0]{figures/zones_study_area.pdf}
   \caption{Map of the different zones (colored polygons) of the 463 km\textsuperscript{2} study area located on the southern plain of Bylot Island, Nunavut Canada. The outline of the snow goose colony area is presented with white dash; we used the outline of the colony in 2017 since it represents an average colony area (74 km\textsuperscript{2}).}
  \label{figure:zones}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth, angle=0]{figures/qarlikturvik_valley.pdf}
  \vspace{-70pt} % Reduce space between figure and caption
  \caption{Intensive nests searching plots (8 km\textsuperscript{2} and 2 km\textsuperscript{2}) located in the Qarlikturvik valley.}
 \label{figure:qarlikturvik_valley}
\end{figure}
\newpage

		\subsubsection*{b. Avian nest monitoring}
Avian nest monitoring was not conducted in 2020 and 2021 due to logistical constraints imposed by the COVID-19 pandemic. Systematic nest monitoring refers here to a systematic sampling approach aimed at documenting all nests within a specified area. Monitoring is considered opportunistic when there is a chance that some nests might not have been detected within a specific area. Nest densities derived from nest sampling could be underestimated due to early nest failure (i.e., failure that happened before our sampling period).

\input{tables/species_year_monitoring.tex}
%label-> table:species_year_monitoring

	\begin{enumerate}[label=\roman*]
	\item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}}\\
	Since 2004, systematic searches of wetland areas have been conducted on the southern side of the glacial river in the Qarlikturvik Valley, and since 2017, in other zones of the study area. This sampling aimed to find all nests of the cackling goose and the glaucous gull. Nest locations of other large wetland-nesting species, including the tundra swan, the red-throated loon and the Pacific loon, were also noted, as these species nest in the same habitat \citep{duchesne2021,gauthier2024a}. Each year, all known or potential nesting sites were revisited. Observers detected nests by walking and scanning around ponds and lakeshores to identify any active nesting sites. These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest. Most of them (geese, swans and gulls) can also reveal their presence with alarm calls or nest defense displays. We are confident that nest detection probability was high for these species given the open landscape.\\

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.5\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
\caption{Number of cackling goose nests found across the study area over time. The red line represents the fitted model over the period 1996-2023 ($y= e^{0.1717x-342.684}$), which was used to estimate the annual abundance of cackling goose between 1996 and 2016. The model presents a strong fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area (R\textsuperscript{2}= 0.74, p= 0.15, n= 5).}
\label{figure:cackling}
\end{figure}

>>>>

\documentclass[a4paper,twoside,12pt]{article}
%% Language %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{scrextend}
\addtokomafont{labelinglabel}{\sffamily}
\usepackage{lmodern} %Type1-font for non-english texts and characters
\usepackage{parskip}
\usepackage{enumitem}
%% Packages for Graphics, Figures & Tables %%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{graphicx} %%For loading graphic files
\usepackage{subcaption}
\usepackage{float}
\usepackage{booktabs} %% for top/midrule
\usepackage{textcomp}
\usepackage[table,xcdraw, dvipsnames]{xcolor}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{makecell}
\usepackage{pdflscape}
%bibliography
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{hyperref}
\hypersetup{
    colorlinks=false,
    pdfborder={0 0 0} % Removes the border
}
\usepackage{geometry}
%% Math Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
%% Line Spacing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
\onehalfspacing       %% 1,5-spacing
\usepackage{placeins}
\usepackage{lineno}
\linenumbers
\usepackage{titlesec}
\titlespacing{\section}{0pt}{*4}{*1.5} 
\titlespacing{\subsection}{20pt}{*4}{*1.5}
\titlespacing{\subsubsection}{40pt}{*4}{*1.5} 
\titlespacing{\paragraph}{60pt}{*4}{*1.5}
\titlespacing{\subparagraph}{80pt}{*4}{*1.5}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begingroup  
  \centering
\textbf{Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
 Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1}, 
 Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]
\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’Études Nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{5} Centre d’Études Nordiques, Centre de Connaissance et d’Exploration de l’Arctique, Musée Canadien de la Nature, Ottawa, ON, Canada\\
\textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
\\[1.5em]
\textbf{Corresponding Author}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
\textit{\href{mailto:joel_bety@uqar.ca}{joel{\textunderscore}bety@uqar.ca}}
\\[1.5em]
\textbf{Open Research statement}:\\
The data set is publicly available at \url{https://datadryad.org/} and the raw data and codes used to extract the data set are publicly available at \url{https://zenodo.org/}.\\
\endgroup
\newpage

\subsection*{Introduction}
 The composition of ecological communities, defined as the abundance of each species within a given community, is fundamental for understanding patterns and processes in community ecology. Variations in community composition can help to detect spatial patterns linked to environmental variations \citep{kemp1990}, assess temporal trends of different groups of species after disturbances \citep{philippi1998, magurran2007}, and understand food web structures \citep{cohen2003}. Additionally, community composition is essential for modeling the dynamics of ecological communities. Dynamic community modelling allows addressing important issues and questions in ecology, such as: determining the relative strength of top-down versus bottom-up forces in communities \citep{krebs2003,legagneux2014}, assessing the ecological resilience of communities under climate change \citep{griffith2019} and evaluating the cascading effects of invasive species in food web \citep{david2017, goto2020}. Dynamic community modelling can also be applied to address practical challenges, including fishery management \citep{plaganyi2007} and the planning of protected areas \citep{okey2004, dahood2020}. 

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.75\linewidth]{figures/food_web.pdf} 
\caption{Synthetic vertebrate food web of the south plain of Bylot Island.}
\label{figure:food_web}
\end{figure}
\subsection*{Objective}
Our main objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. Our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments. Additionally, we aim to provide the average body mass for each species in the community, enabling the conversion of abundances into biomasses.
\section*{\textcolor{Blue}{Class I. Data Set Descriptors}}
    \section*{A. Data set identity} Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut
    \section*{B. Data set identification codes}
     \texttt{BYLOT-species\_taxonomy.csv}\\
     \texttt{BYLOT-species\_abundance.csv}\\
     \texttt{BYLOT-mean\_species\_abundance.csv}\\
     \texttt{BYLOT-mean\_species\_body\_mass.csv}\\
     \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\

    \section*{C. Data set description}
                \subsection*{1. Originators}
                \textbf{Gilles Gauthier}, Centre d’Études Nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
                \textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada

                \subsection*{2. Abstract}
Arctic ecosystems present unique opportunities for community-wide monitoring, in part due to their relatively low species richness. However, conducting research in these remote environments poses significant logistical challenges, resulting in long-term monitoring being exceedingly rare. Here, we focus on the long-term, intensive ecological monitoring efforts conducted on the south plain of Bylot Island (almost 400 km\textsuperscript{2}, Nunavut, Canada), which has generated a remarkable dataset spanning up to 30 years, a rarity in tundra ecosystems. Our goal is to synthesize this dataset and upscale vertebrate abundance data at the landscape level, a prerequisite to conduct community-level analyses. We have standardized data obtained with different field methods to provide readily usable long-term time series of abundance for 35 vertebrate species (30 birds and 5 mammals) present in the study system. Monitoring data includes intensive capture-mark-recapture density estimates of lemmings on trapping grids, systematic or opportunistic nest monitoring conducted across the entire study area or within specific plots for all bird species, transects of vertebrate counts distributed throughout the study area, daily incidental observations of vertebrates and satellite tracking of fox movements. Annual abundance of species was estimated at the landscape level, accounting for spatial variations. Furthermore, we provide body masses for each species, derived from empirical onsite measurements for 18 species and from the literature for the remaining species. Body mass is essential to convert species abundance into biomass for studies of trophic fluxes and ecosystem processes. Our dataset provides a unique opportunity for holistic empirical studies of ecological communities, allowing a deeper understanding of community structure and dynamics. Considering that the study site is a pristine and protected area that has experienced minimal anthropogenic impact, it can also provide an ideal baseline for investigating the impacts of global changes on high-latitude terrestrial ecosystems.
   \section*{D. Key words/phrases}
 Bylot Island, Canadian Arctic, Arctic tundra, 1993-2023, long-term monitoring, biodiversity monitoring, community composition, species abundance, species density, species biomass, species body mass, food web
  \newpage

\section*{\textcolor{Blue}{Class II. Research origin descriptors}}
    \section*{A. Overall project description}
    \subsection*{1. Identity} Understanding the structure and dynamics of Arctic terrestrial vertebrate communities
        \subsection*{2. Originators:}   
        \textbf{Gilles Gauthier}, Centre d’études nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
                \textbf{Joël Bêty},  Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Département de biologie, chimie et géographie, Université du Québec à Rimouski, Rimouski, QC, Canada \\
                \textbf{Pierre Legagneux}, Chaire de Recherche Sentinelle Nord sur l’impact des migrations animales sur les écosystèmes nordiques, Centre d’Études Nordiques, Centre de la science de la biodiversité du Québec, Centre d’Études Biologiques de Chizé (CEBC-CNRS) Département de Biologie, Université Laval, Québec, QC, Canada

@@

    \section*{B. Specific subproject description}
        \subsection*{1. Site description}
                        \subsubsection*{a. Site type}
  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic. An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:table_species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:table_species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (\textit{Vulpes vulpes}) was ambiguous. While the presence of breeding pairs has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare. Therefore, we decided to exclude this species.
                        \input{tables/table_species_name_strategy.tex}
%label-> table:table_species_name_strategy
\newpage
                \subsubsection*{b. Geography} Our 463 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
                \subsubsection*{c. Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
                \subsubsection*{d. Geology} See \citet{klassen1993} for a complete and detailed description of the geology of the study area.
                \subsubsection*{e. Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
                \subsubsection*{f. Site history} See \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
                \subsubsection*{g. Climate} The mean temperature in July is 6°C, and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}. The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds. \citep{gauthier2024a}.
 \newpage

        \subsection*{2. Experimental or sampling design}
                        \subsubsection*{a. Permanent plots}
The study area is divided into 9 zones based on the sampling method and the level of field effort applied in each zone (\textbf{Figure \ref{figure:zones}}). Long-term monitoring of the community began in the 1990s in the Qarlikturvik valley \citep{gauthier2013, gauthier2024a}, which represents the zone of the study area with the highest annual sampling effort. Within the Qarlikturvik valley, the sampling is concentrated on the southern side of the glacial river (\textbf{Figure \ref{figure:qarlikturvik_valley}}), where the main research infrastructure is located. Another zone with extensive sampling efforts is Camp 2, located at the core of the snow goose colony, where the primary focus is to monitor snow goose nests. However, nests of many other avian species are also monitored within and around the colony in this zone. Camp 3, Pointe Dufour, Goose Point, and Malaview are zones where intensive sampling efforts are conducted annually, albeit for a relatively brief period (approximately one week) during the breeding season of most species \citep{gauthier2024a}. The upland zones in the study area (defined as areas approximately 300 meters above sea level or more) are the Black Plateau, Southern Plateau, and Camp 3 Plateau. These zones are primarily visited to assess raptor nesting activity \citep{beardsell2016}. The zone between the Qarlikturvik valley and Camp 3 received very little sampling effort and is therefore excluded from the study area.
\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth, angle=0]{figures/zones_study_area.pdf}
   \caption{Map of the different zones (colored polygons) of the 463 km\textsuperscript{2} study area located on the southern plain of Bylot Island, Nunavut Canada. The outline of the snow goose colony area is presented with white dash; we used the outline of the colony in 2017 since it represents an average colony area (74 km\textsuperscript{2}).}
  \label{figure:zones}
\end{figure}
\begin{figure}
\centering
  \includegraphics[width=0.8\textwidth, angle=0]{figures/qarlikturvik_valley.pdf}
  \vspace{-70pt} % Reduce space between figure and caption
  \caption{Intensive nests searching plots (8 km\textsuperscript{2} and 2 km\textsuperscript{2}) located in the Qarlikturvik valley.}
 \label{figure:qarlikturvik_valley}
\end{figure}
                \subsubsection*{b. Avian nest monitoring}
Avian nest monitoring was not conducted in 2020 and 2021 due to logistical constraints imposed by the COVID-19 pandemic. Systematic nest monitoring refers here to a systematic sampling approach aimed at documenting all nests within a specified area. Monitoring is considered opportunistic when there is a chance that some nests might not have been detected within a specific area. Nest densities derived from nest sampling could be underestimated due to early nest failure (i.e., failure that happened before our sampling period).
\input{tables/table_species_year_monitoring.tex}
%label-> table:table_species_year_monitoring
        \begin{enumerate}[label=\roman*]
        \item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}}\\
        Since 2004, systematic searches of wetland areas have been conducted on the southern side of the glacial river in the Qarlikturvik Valley, and since 2017, in other zones of the study area. This sampling aimed to find all nests of the cackling goose and the glaucous gull. Nest locations of other large wetland-nesting species, including the tundra swan, the red-throated loon and the Pacific loon, were also noted, as these species nest in the same habitat \citep{duchesne2021,gauthier2024a}. Each year, all known or potential nesting sites were revisited. Observers detected nests by walking and scanning around ponds and lakeshores to identify any active nesting sites. These large species can be seen from a relatively long distance sitting on the nest or when flushing from the nest. Most of them (geese, swans and gulls) can also reveal their presence with alarm calls or nest defense displays. We are confident that nest detection probability was high for these species given the open landscape.\\

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.5\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
\caption{Number of cackling goose nests found across the study area over time. The red line represents the fitted model over the period 1996-2023 ($y= e^{0.1717x-342.684}$), which was used to estimate the annual abundance of cackling goose between 1996 and 2016. The model presents a strong fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area (R\textsuperscript{2}= 0.74, p= 0.15, n= 5).}
\label{figure:cackling}
\end{figure}
